### 2022-10-19 latex/supplementary.tex
Add files via upload

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,mathtools,setspace,enumitem,epsfig,titlesec,verbatim,color,array,multirow,comment,graphicx,hyperref,blkarray}
%\usepackage[sort&compress]{natbib} % ProcB
\usepackage[super,sort&compress,comma]{natbib} % NComms
\usepackage[bf,small]{caption}
\usepackage[export]{adjustbox}
\usepackage[top=2.5cm,left=2.8cm,right=2.8cm,bottom=3.2cm]{geometry} 
\smallskip

\definecolor{darkred}{rgb}{0.6,0,0}
\definecolor{darkblue}{rgb}{0,0.3,0.8}

\newcommand{\christian}[1]{\textcolor{blue}{{\bf CH:} #1}}

\titleformat{\section}{\sffamily \fontsize{12}{20}\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\sffamily \fontsize{11}{20}\bfseries}{\thesubsection}{1em}{}

\renewcommand{\figurename}{Supplementary Figure}

\newcommand{\FigIllustration}{{\bf Fig.~1}}

\newtheoremstyle{plainCl1}% name
{9pt}%      Space above, empty = 'usual value'
{15pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{.}%        Punctuation after thm head
{2mm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\newtheoremstyle{plainCl2}% name
{9pt}%      Space above, empty = 'usual value'
{15pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{.}%        Punctuation after thm head
{4mm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\theoremstyle{plainCl1}
\newtheorem{theorem}{Theorem}
\newtheorem{Prop}{Proposition}

\theoremstyle{plainCl2}
\newtheorem{lemma}{Lemma}
\newtheorem{Cor}{Corollary}

\newcommand{\ALLD}{\emph{D}}

\newcommand{\ai}{\mathbf{A}^i}

\title{\sffamily \Large Supplementary Information\\[0.1cm] {\bfseries Introspection dynamics in asymmetric multiplayers games}}
\date{\empty}
\author{\parbox[c]{16cm}{\centering \onehalfspacing \fontsize{11}{12}\selectfont Marta Couto$^1$ and Saptarshi Pal$^1$\\[0.2cm]
$^1$Max Planck Research Group Dynamics of Social Behavior, Max Planck Institute for Evolutionary Biology, 24306~Ploen, Germany}}

\begin{document}
\maketitle
\onehalfspacing
\section*{The general model for introspection dynamics}
We consider an asymmetric normal form game with $N (>2)$ players. Each player, $i$, has access to their action set, $\ai$, in which there are $m_i$ possible actions that they can play, $\ai = \{a^i_i, a^i_2, ..., a^i_{m_i}\}$. The payoff for an individual $i$ depends on what everyone plays and is represented by $\pi^i(a^1, a^2, ..., a^N)$, where $(a^1, a^2,...,a^N) \in \mathbf{A}^1 \times \mathbf{A}^2 \times ... \times \mathbf{A}^N$. In this model we only consider pure strategies for players. So, player $i$ has only $m_i$ possible strategies.

\noindent The players update their strategies over time using the introspection dynamics \cite{couto2022introspection}.

\section*{Examples of multiplayer games and results}

\section*{Supplementary References}
\bibliographystyle{unsrt}
\bibliography{bibliography.bib }
\end{document}